---
title: "OSUN R Users Community"
subtitle: "Kevin O'Brien"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

```{r warning=FALSE,message=FALSE,echo=FALSE}
library(tidyverse)
library(magrittr)
library(faraway)
data(cheddar)


Fit_1 <- lm(taste ~ Acetic + Lactic, data = cheddar)
Fit_2 <- lm(taste ~ Acetic + H2S, data = cheddar)
Fit_3 <- lm(taste ~ H2S + Lactic, data = cheddar)

Fit_4 <- lm(taste ~ Acetic + H2S + Lactic, data = cheddar)
```

---

### Recall: Assumption of Constant Variance

#### Homoscedasticity

* ***Homoscedascity*** is the technical term to describe the variance of the residuals being constant across the range of predicted values. 

* ***Heteroscedascity*** is the converse scenario : the variance differs along the range of values.



#### Diagnostic Plots

Suppose you plot the individual residuals against the predicted value, the variance of the residuals predicted value should be constant (Diagnostic Plot 1). 



```{r include=FALSE,echo=FALSE}

fit <- Fit_4
```

---

### {car} R Package

Cheeses Data

```{r warning=FALSE,message=FALSE}	

# Evaluate homoscedasticity

# non-constant error variance test

library(car)
ncvTest(fit)
```

---

### +residual plots


* component+residual plots, also called partial-residual plots, for linear and generalized linear models.

* a graphical technique that attempts to show the relationship between a given independent variable and the response variable given that other independent variables are also in the model.

* Ceres plots are a generalization of component+residual (partial residual) plots that are less prone to leakage of nonlinearity among the predictors. 

---

### Nonlinearity

#### component + residual plot 

```{r out.width = "70%" }	

crPlots(fit)
```

---

### Nonlinearity

#### ceres plot

```{r out.width = "70%" }

ceresPlots(fit)
```

---

### Nonlinearity


Test for Autocorrelated Errors

```{r}

durbinWatsonTest(fit)

```

---

### avplot & av plots

* The avplot command is short for added-variable plot and draws exactly that. 

* The other name for added variable plot is a partial-regression leverage plot. 
