---
title: "OSUN R Users Community"
subtitle: "Kevin O'Brien"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

```{r warning=FALSE,message=FALSE,echo=FALSE}
library(tidyverse)
library(magrittr)
library(faraway)
data(cheddar)


Fit_1 <- lm(taste ~ Acetic + Lactic, data = cheddar)
Fit_2 <- lm(taste ~ Acetic + H2S, data = cheddar)
Fit_3 <- lm(taste ~ H2S + Lactic, data = cheddar)

Fit_4 <- lm(taste ~ Acetic + H2S + Lactic, data = cheddar)
```

Diagnostic Plots for Linear Models with R
===================================================

### Plot Diagnostics for an <tt>lm()</tt> Object

There are six plots (selectable by <tt>which=</tt>) are currently available: 

*  a plot of residuals against fitted values, 
*  a Normal Q-Q plot, 
*  a Scale-Location plot of <tt>sqrt( $\mid \mbox{residuals} \mid$ })</tt> against fitted values, 

*  a plot of Cook's distances versus row labels, 
*  a plot of residuals against leverages, 
*  a plot of Cook's distances against *leverage/(1-leverage)*.

By default, the first three and 5 are provided, if you just type something like <tt?plot(fit)</tt>.

---

### Diagnostic Plot 1

* The first one displays the residuals vs. the fitted values we use this to evlauate the mean, variance and correlation of residuals.

* If our assumptions of constant variance and uncorrelated residuals are violated we **may** be able to correct this with a variance-stabilizing transformation.

* see <tt>car::ncevTest()</tt>


---
### Diagnostic Plot 1

```{r eval=FALSE}
  plot(Fit_4,
     which=1,
     pch=16,lwd=1.2)
```  

Just increment the "<tt>which=</tt>" argument with any integer between 1 and 6


---
### Diagnostic Plot 1

```{r out.width = "80%",echo=FALSE }
plot(Fit_4,which=1,
     pch=16,lwd=1.2)
```


---
### Diagnostic Plot 2

* The second plot helps us check the normality of the residuals. 

* If the residuals are indeed normal, they should fall along the dashed line.

* Remember that the normality assumption for our errors allows us to determine
the standard errors of our coefficients and predictions.

---
### Diagnostic Plot 2



```{r echo = FALSE,out.width = "80%" }
plot(Fit_4,which=2,
     pch=16,lwd=1.2)

```

---
### Diagnostic Plots 3

* The ***Scale-Location*** plot, also called 'Spread-Location' (or 'S-L' plot), takes the square root of the absolute residuals in order to diminish skewness (**sqrt($\mid E\mid$)**) is much less skewed than $\mid E\mid$ for Gaussian zero-mean E).
 
 
---
### Diagnostic Plots 3

```{r echo = FALSE,out.width = "80%" }
plot(Fit_4,which=3,
     pch=16,lwd=1.2)

```


---
### Diagnostic Plots 4

* This plot details the Cook's Distance for each observation. 
* We will revert to this later.
 
---

### Diagnostic Plots 4

```{r echo = FALSE,out.width = "80%" }
plot(Fit_4,which=4,
     pch=16,lwd=1.2)

```

--- 

### Extracting Cook's Distance

```{r}
cooks.distance(Fit_4) %>% 
  head(5) %>% 
  round(4) %>%
  t()
```

---

### Diagnostic Plots 5

* The ***Residual-Leverage*** plot shows contours of equal Cook's distance, for values of ***cook.levels*** (by default 0.5 and 1) and omits cases with leverage one with a warning. 

* If the leverages are constant the plot uses factor level combinations instead of the leverages for the x-axis. 

* **(The factor levels are ordered by mean fitted value.)**

---

### Diagnostic Plots 5


```{r echo = FALSE,out.width = "80%" }
plot(Fit_4,which=5,
     pch=16,lwd=1.2)

```

---

### Diagnostic Plots 6

* The final plot will display our residuals vs. their leverage. 
* The dashed red lines are level curves that denote a particular value of Cook's distance.
* We will pay attention to points lying beyond the distance of 1. 
* Notice that when we have data with row labels, the points will be labeled with their names. Otherwise, the row number will be shown.


---

### Diagnostic Plots 6

```{r echo = FALSE,out.width = "80%" }
plot(Fit_4,which=6,
     pch=16,lwd=1.2)

```

---