---
title: "OSUN R Users Community"
subtitle: "Kevin O'Brien"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

### Linear modelling with R (Cheeses)


#### Cheddar Cheese taste

* As cheese ages, various chemical processes take place that determine the taste of the final product. 

* This dataset contains concentrations of various chemicals in 30 samples of mature cheddar cheese, and a subjective measure of taste for each sample. 

* The variables "Acetic" and "H2S" are the natural logarithm of the concentration of acetic asid and hydrogen sulfide respectively. 

* The variable "Lactic" has not been transformed.

#### Reference: 

* Moore, David S., and George P. McCabe (1989). Introduction to the Practice of Statistics.

---

### Linear modelling with R (Cheeses)


* Number of cases: 30

#### Variable Names:

* **Case**: Sample number
* **taste**: Subjective taste test score, obtained by combining the scores of several tasters
* **Acetic**: Natural log of concentration of acetic acid
* **H2S**: Natural log of concentration of hydrogen sulfide
* **Lactic**: Concentration of lactic acid


---

### Linear modelling with R (Cheeses)


```{r warning=FALSE,message=FALSE}
library(tidyverse)
library(magrittr)
library(faraway)
data(cheddar)


head(cheddar)
```


---

### Linear modelling with R (Cheeses)


```{r}
Fit_1 <- lm(taste ~ Acetic + Lactic, data = cheddar)
Fit_2 <- lm(taste ~ Acetic + H2S, data = cheddar)
Fit_3 <- lm(taste ~ H2S + Lactic, data = cheddar)

Fit_4 <- lm(taste ~ Acetic + H2S + Lactic, data = cheddar)
```

---

### Linear modelling with R (Cheeses)


#### Aikaike Information Criterion

```{r}
AIC(Fit_1)

AIC(Fit_2)

AIC(Fit_3)

AIC(Fit_4)

```

---

### {modelr}

Compute model quality for a given dataset

Three summaries are immediately interpretible on the scale of the response variable:

* <tt>rmse()</tt> is the root-mean-squared-error
* <tt>mae()</tt> is the mean absolute error
* <tt>qae()</tt> is quantiles of absolute error.

---

### {modelr}

#### Root Mean Square Error

```{r}
library(modelr)
rmse(Fit_4,cheddar)
```


---

### {modelr}

#### mean absolute error

```{r}
mae(Fit_4,cheddar)
```

---

### {modelr}

```{r}
qae(Fit_4,cheddar)
```

---

### {modelr}

Other summaries

* <tt>mape()</tt> mean absolute percentage error.
* <tt>rsae()</tt> is the relative sum of absolute errors.
* <tt>mse()</tt> is the mean-squared-error.
* <tt>rsquare()</tt> is the variance of the predictions divided by the variance of the response.

---

### {modelr}

```{r}
rsquare(Fit_4,cheddar)
```
