\section*{Exercise 20}
 
Generate a normally distributed random sample of size 300. What is the mean and standard deviation of this sample? What are the 0.9, 0.95, 0.975 and 0.99 quantiles quantiles of this sample? Compare both sets of results.
Use R to determine the 90th, 95th and 99th quantiles of a normal distribution, with mean 100 and standard deviation 15.
 
%----------------------------------------------------------------%
\section*{Exercise 21}
IQ scores are assumed to have a normal distribution with mean 100 and standard deviation 15. 
 
\begin{itemize}
\item What IQ would you have if you were in the 80th percentile?
\item Estimate the threshold for the top 10 percent?
\item What is the probability of having an IQ above 142?
\item What is the probability of having an IQ below 97?
\end{itemize}
 
 %----------------------------------------------------------------%
\section*{Exercise 22}
 
Use the uniform distribution to simulate 100 throws of two dice. The outcome is the combined values of both dice. Use the appropriate R command to discretize values.
\begin{itemize}
\item  What is the mean and standard deviation of the outcomes?
\item  Make a stem-and-leaf plot of the outcomes.
\item Make a histogram of the outcomes. (hint: use breaks =seq(1.5,12.5))
\end{itemize}
 
%----------------------------------------------------------------%
 
\section*{Exercise 23}

Add up all the numbers for 1 to 50
\begin{itemize}
\item  Using the sum() command.
\item  using a ‘for’ loop.
\end{itemize}
